---
title: "Sensors_Functions"
author: "Shelmith Kariuki"
date: "July 26, 2017"
output: html_document
---

#This RMarkdown document contains codes that are used to generate the following functions:

##1. sensor_raw       : Reads in the raw data and returns the payloads.
##2. probe_location   : Extracts and returns location probe data from the payloads.
##3. probe_sms        : Extracts and returns sms messages probe data from the payloads.
##4. probe_calls      : Extracts and returns call log probe data from the payloads.
##5. probe_event      : Extracts and returns communication event probe data from the payloads.
##6. sensor_population: Takes in each probe's data and identifies the users whose information                                             available for that  particular probe.The function then reads-in the master dataset(one                            with the list of all the participants). A new variable is then generated, for each sensor,                         that indicates whether we have data for each participant in the master dataset. This                              variable is binary: Available and Not Available.
##7.count_commtype    : Displays the count for each communication type, Phones and SMS.
##8.stats_phone       : Takes in the communication event probe, filters all communications done via                                       phone and finally counts the daily number, and average number of incoming, outgoing and                           missed calls per user.
##9.stats_sms         : Takes in the communication event probe, filters all communications done via                                       sms and finally counts the daily number, and average number of incoming and outgoing sms                          messages per user
##10.ratio_phone      : Displays the ratio of outgoing calls to incoming calls.
##11.ratio_sms        : Displays the ratio of outgoing sms messages to incoming calls.



#sets the working directory
```{r setup, include=FALSE}
basedir <- "C:\\Users\\Shelmith\\Documents\\Projects\\Sensors\\Data"
knitr::opts_knit$set(root.dir = basedir)
```

#loading the libraries

```{r}
library(readr)
library(dplyr)
library(tidyjson)
library(lubridate)
library(tidyr)
```

##sensor_raw function

```{r}

sensor_raw<-function(raw_data){
              pl <- raw_data %>% select(user_id, payload) %>% collect %>% 
                  mutate(payload = paste("[", payload, "]", sep = "")) %>% 
                    as.tbl_json(.,json.column="payload") %>% gather_array  %>% 
                        spread_values(
                          probe=jstring("PROBE"),
                          timestamp=jnumber("TIMESTAMP"),
                          guid=jstring("GUID"))
                           (plTypes <- pl %>% distinct(probe)) 
return(pl)
}
suppressMessages(psRaw <- read_csv(file =".\\RawData\\sensorinternal12072017_01.csv"))
sensor_raw(psRaw)
```

##probe_location

```{r}
probe_location<-function(pl){
                    filterPayload <- function(pl, probeFilter) {
                        pl %>% filter(grepl(probeFilter, probe))%>%select(-probe)
                      }
                    probeFilter <- "\\.LocationProbe$"
                     tryCatch({
                          pl %>% sampleProbe(probeFilter) 
                        }, error=function(e){
                            print("LocationSensorProbe data is not available")
                       })
                    locationProbe <- pl %>% filterPayload(probeFilter) %>% 
                         spread_values(
                          gpsAvailable=jlogical("GPS_AVAILABLE"),
                          latitude=jnumber("LATITUDE"),
                          timeFix=jstring("TIME_FIX"),
                          longitude=jnumber("LONGITUDE"),
                          networkAvailable=jlogical("NETWORK_AVAILABLE"),
                          accuracy=jnumber("ACCURACY"),
                          provider=jstring("NETWORK"))%>% 
                            select(user_id, timestamp, gpsAvailable, latitude, longitude, networkAvailable, accuracy) %>%                                 mutate(event_timestamp = as.POSIXct(timestamp, origin="1970-01-01")) %>% 
                                     mutate(event_hour = hour(event_timestamp),event_date = as.Date(event_timestamp)) 
  
return(locationProbe)
}
probe_location(sensor_raw(psRaw))
```

##probe_sms function takes in the payloads and extracts the smsmessages probe.

```{r}
probe_sms<-function(pl){
  
  filterPayload <- function(pl, probeFilter) {
  pl %>% filter(grepl(probeFilter, probe))%>%select(-probe)
  }
communicationLogProbesSmsMessages <- pl %>% filterPayload("\\.CommunicationLogProbe$")%>%
   enter_object("SMS_MESSAGES") %>% gather_array %>%
   spread_values(
     numberName=jstring("NUMBER_NAME"),
     normalizedHash=jstring("NORMALIZED_HASH"),
     messageDirection=jstring("MESSAGE_DIRECTION"),
     number=jstring("NUMBER"),
     messageTimestamp=jnumber("MESSAGE_TIMESTAMP")
   ) %>% select(-array.index)%>% 
  select(user_id, numberName, messageDirection, number, messageTimestamp) %>% 
  mutate(msg_timestamp = as.POSIXct(messageTimestamp/1000, origin="1970-01-01")) %>%
  mutate(event_hour = hour(msg_timestamp),
         event_date = as.Date(msg_timestamp))
return(communicationLogProbesSmsMessages)
}
probe_sms(sensor_raw(psRaw))
 
```

##probe_calls function takes in the payloads and extracts the call log probe.

```{r}
probe_calls<-function(pl){
  
 filterPayload <- function(pl, probeFilter) {
  pl %>% filter(grepl(probeFilter, probe))%>%select(-probe)
 }

  communicationLogProbesPhoneCalls <- pl %>% filterPayload("\\.CommunicationLogProbe$")%>%
   enter_object("PHONE_CALLS") %>% gather_array %>%
   spread_values(
     numberName=jstring("NUMBER_NAME"),
     callDuration=jnumber("CALL_DURATION"),
     numberLabel=jstring("NUMBER_LABEL"),
     normalizedHash=jstring("NORMALIZED_HASH"),
     numberType=jstring("NUMBER_TYPE"),
     callTimestamp=jnumber("CALL_TIMESTAMP"),
     number=jstring("NUMBER")
   ) %>% select(-array.index)%>% 
  select(user_id, numberName, callDuration, numberLabel, numberType, number, callTimestamp) %>% 
  mutate(call_timestamp = as.POSIXct(callTimestamp/1000, origin="1970-01-01")) %>%
  mutate(event_hour = hour(call_timestamp),
         event_date = as.Date(call_timestamp))
 return(communicationLogProbesPhoneCalls)
}
probe_calls(sensor_raw(psRaw))
```

##probe_event function takes in the payloads and extracts the communication event probe.

```{r}
probe_event<-function(pl){
  filterPayload <- function(pl, probeFilter) {
  pl %>% filter(grepl(probeFilter, probe))%>%select(-probe)
  }
  probeFilter <- "\\.CommunicationEventProbe$"
CommunicationEventProbe <- pl %>% filterPayload(probeFilter) %>% 
   spread_values(
     name=jstring("NAME"),
     timestamp=jnumber("TIMESTAMP"),
     comm_timestamp=jstring("COMM_TIMESTAMP"),
     probe=jstring("PROBE"),
     comm_direction=jstring("COMMUNICATION_DIRECTION"),
     number=jstring("NUMBER"),
     communication_type=jstring("COMMUNICATION_TYPE")
    )%>% 
  select(user_id, comm_timestamp, comm_direction, number, communication_type) %>% 
  mutate(event_timestamp = as.POSIXct(as.numeric(comm_timestamp)/1000, origin="1970-01-01")) %>%
  mutate(event_hour = hour(event_timestamp),
         event_date = as.Date(event_timestamp))
return(CommunicationEventProbe)
}
probe_event(sensor_raw(psRaw))

```

##sensor_population function

```{r}
sensor_population<-function(master_data,location_sensor,calls_sensor,sms_sensor,commevent_sensor ){
unique_registered<-master_data%>%
  filter(!duplicated(user_id))%>%
  mutate(location=ifelse(user_id%in%unique(location_sensor$user_id),"Available","NotAvailable"),
            calls=ifelse(user_id%in%unique(calls_sensor$user_id),"Available","NotAvailable"),
              sms=ifelse(user_id%in%unique(sms_sensor$user_id),"Available","NotAvailable"),
          event =ifelse(user_id%in%unique(commevent_sensor$user_id),"Available","NotAvailable"))%>%
  select(-email)
return(unique_registered)
}
#Reading in the users register
master_data<-read.csv("C:\\Users\\Shelmith\\Documents\\Projects\\Sensors\\Data\\SupplementaryData\\registered_users.csv")
sensor_population(master_data,probe_location(sensor_raw(psRaw)),probe_calls(sensor_raw(psRaw)),
                  probe_sms(sensor_raw(psRaw)),probe_event(sensor_raw(psRaw)))
```

##count_commtype function 

```{r}
count_commtype<- function(df){
  commtype_count<-df%>% 
  group_by(user_id,event_date)%>%
  count(communication_type)%>%
  spread(communication_type, n)
  return(commtype_count)
}
count_commtype(probe_event(sensor_raw(psRaw)))

```

##stats_phone function 

```{r}
#df is the comm event sensor data.
stats_phone<-function(df){
  phones_daily<- df %>% 
      filter(communication_type == "PHONE")%>%
          group_by(user_id,event_date)%>%
              count(comm_direction)%>%
                  mutate(average_no.calls = ceiling(mean(n)))%>%
                        spread(comm_direction, n)
                      
return(phones_daily)
}
stats_phone(probe_event(sensor_raw(psRaw)))

```

##stats_sms function 

```{r}
#df is the comm event sensor data.
stats_sms<-function(df){
  sms_daily<- df %>% 
      filter(communication_type == "SMS")%>%
          group_by(user_id,event_date)%>%
              count(comm_direction)%>%
                  mutate(average_no.messages = ceiling(mean(n)))%>%
                        spread(comm_direction, n)
                      
return(sms_daily)
}
stats_sms(probe_event(sensor_raw(psRaw)))
```

##ratio_phone function displays the ratio of outgoing calls to incoming calls.

```{r}
ratio_phone<-function(df){
phone_ratio<-df %>% 
      filter(communication_type == "PHONE")%>%
          group_by(user_id,event_date)%>%
              count(comm_direction)%>%
                  spread(comm_direction, n)%>%
                        mutate(ratio= OUTGOING / INCOMING)  
return(phone_ratio)
}
ratio_phone(probe_event(sensor_raw(psRaw))) 
```

##ratio_sms function displays the ratio of outgoing sms messages to incoming calls.

```{r}
ratio_sms<-function(df){
sms_ratio<-df %>% 
      filter(communication_type == "SMS")%>%
          group_by(user_id,event_date)%>%
              count(comm_direction)%>%
                  spread(comm_direction, n)%>%
                        mutate(ratio= OUTGOING / INCOMING)  
return(sms_ratio)
}
ratio_sms(probe_event(sensor_raw(psRaw))) 
```

##End of RMarkdown



